---
title: "Task 8"
author: "Patrick McKenzie"
date: "March 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include = FALSE}
library(dplyr)
library(nycflights13)
```

# Resources
* Advanced R chapters 6, 9 - 11
* R for Data Science chapters 14-17
* [This blog post](http://mazamascience.com/WorkingWithData/?p=912) about the `tryCatch() function`

# Tasks

# Function tasks
* Complete this task by creating an R notebook (.Rmd file) and accompanying .md file
## Functions and error handling:
Write a function that fulfills the following criteria:
* It should be tidyverse compatible (i.e., the first argument must be a data frame)
* It should add two arbitrary columns of the data frame (specified by the user) and put them in a new column of that data frame, with the name of the new column specified by the user
* It should throw an informative warning if any invalid arguments are provided. Invalid arguments might include:
    * The first argument is not a data frame
    * Less than two valid columns are specified to add (e.g., one or both of the column names isn't in the supplied data frame) 
    * The columns specified are not numeric, and therefore can't be added - use `tryCatch()` for this
* If the columns to add aren't valid but the new column name is, the function should create a column of NA values
```{r}
add_two_columns <- function(data, column.headings, new.column.name = NULL) {
if (is.data.frame(data) == FALSE) {
  stop("First argument must be a data frame.")
}
if (length(column.headings) != 2) {
  stop("Argument that is a vector of two column names must be provided.")
}
if (is.tbl(data)) {
  stop("Your data is a tbl! Don't use tbls -- just use as.data.frame() and make everything easier.")
}
col1 <- data[,column.headings[1]]
col2 <- data[,column.headings[2]]
if (is.numeric(col1) && is.numeric(col2)) {
  newcol <- col1  + col2
  dim(newcol) <- c(length(newcol),1)
  if (length(new.column.name) == 1) {
    colnames(newcol) <- new.column.name
  }
  else {
    colnames(newcol) <- "new.column"
  }
  cbind.data.frame(data,newcol)
}
else {
  if (length(new.column.name == 1)) {
    warning("Columns specified are not numeric, returning column of NA values.")
    na.column <- rep(NA, nrow(data))
    dim(na.column) <- c(length(na.column),1)
    colnames(na.column) <- new.column.name
    cbind.data.frame(data,na.column)
  }
  else {
    stop("Columns specified are not numeric and column name not provided.")
  }
}
}
exampledf <- as.data.frame(nycflights13::flights)
example <- add_two_columns(exampledf,c("air_time","distance"),"my.new.column")
head(example)
```

## Loop and performance metric tasks
* Write a function named that uses a for loop to calculate the sum of the elements of a vector, which is passed as an argument (i.e., it should do the same thing that `sum()` does with vectors). `your_fun(1:10^4)` should return 50005000.
* Use the `microbenchmark::microbenchmark` function to compare the performace of your function to that of `sum` in adding up the elements of the vector `1:10^4`. The benchmarking code should look something like:
```
test.vec <- 1:10^4
microbenchmark(
    my_sum(test.vec),
    sum(test.vec)
    )
```
Is there a difference? Why?

    